NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-06-18 paper/itemizedReviewResponse.tex

R1(3):
+ interesting topic
+ usage scenarios are interesting
- description of NLP NN and attention models are not sufficiently detailed in the paper
- details of collaboration with domain experts are missing
- justification of tasks are not sufficiently detailed
- differences between graph and matrix view should be explained in more detail
- density contour in the prediction view needs to be explained in more detail
- claims about extension of visualization system to non-NLI tasks is not supported
- need to clarify difference between proposed perturbation-driven analysis and gradient-based analysis
- qualitative evaluation is weak

R2 (5):
+ important area for visualization research
+ article is well-written
+ application with linked views is well-designed and powerful
+ system supports debugging of inference models
- scalability of system is not discussed

R3 (2.5):
- article is dense and difficult to read
- rationale for design choices are not clear
- core contributions are not highlighted, too much emphasis on background material (explanation of NLI), not enough on the visualization system itself
- claims based on user evaluation aren't supported

Summary of major changes:
1. Revise section 6 to better convey the relationship between the different senarios and provide further clarification on the result observered in 6.3

@@

------------------------------- Reviewer 1 -------------------------------

Q1. The usage scenarios also look interesting, especially the Scenario 3. Does the result to some extent suggest that the attention mechanism is not a very efficient component in this task? I expect the authors dig it further.

A: To interpret the result, we need to carry out a case by case analysis, as the separation of the stages depends on the individual implementation. In this case, the result indicates that the tunable parameters of the attention stage (the transformation before the alignment operation, see details in Appendix A) has less impact on the final prediction compared to the other two stages. The domain experts comment that this observation is interesting and worth further investigation from the perspective of NLP research. However, this result does not necessarily imply the limitation of attention mechanism as the core alignment computation does not have tunable parameters that can be modified by the proposed optimization. We revised section 6.3 to clarify this point.

@@

Q2. In Section 2, the neural net in NLP and attention mechanisms are introduced. But the descriptions are unsatisfactory. There is no formal definition or description of the model and the attention mechanism. It's still unclear how the attention mechanism is applied in the model. I would suggest the authors include a concise and formal introduction in the background part and include some formula.

A: The exact computation of attention can vary from model to model, though the overall concept is shared as they are specially designed for understanding the corresponding relationship between sequence of texts.

@@

Q3, In Section 4, the paper indicates that the system is developed under a long-term collaboration with NLP experts, but there is no description describing the details of the collaboration (the number of experts, the length and the form of the collaboration, the frequency of the meetings). It's not until Section 7 did I learn that there are 2 NLP experts in collaboration. The task analysis can also be improved by including more detailed justification of the three tasks.

A: We had added additional details of the collobration in the begining of the section 4, include the number of NLP collaborators, frequency, duration of the meeting for this specific tasks.

@@

Q4, The presentation of the system design in Section 5 could be improved.
- In Section 5.2, Attention View include both a bipartite graph and a matrix view, which visualize the same information. From the description of the paper, the graph view is only good for highlighting the most dominant alignments. I suspect that matrix view can also reveal the most dominant alignments by identifying the most salient color. Please justify the design decision.

A: The advantage of the bipartite graph is not only the ability to allow the domain experts quickly understand the domain alignment but more importantly the corresponding words. However, in the matrix view, it is much harder to instantaously link the alignment with the corresponding words. [We update the text to better reflect this point]

Q5, It's unclear how the simplification of attention representation is done from Section 5.2, please include more details.

A: The simplification is achieve by collapsing the dependency tree structure. We assume by focusing on the more important grammar structure, we can remove the decorative words and the reduce sentence length to focus on the more crucial part of the sentences.

@@

Q6, In Section 5.3, how do you compute the density contour? Why is a contour needed in the analysis?

A: We employed the standard kernel density estimation technique in 2D. As stated in Section 5.3, the primary motivation is to highlight overlaped point (where multiple prediction fall on top of each other) and identify outliers.

Q7, It's better if the authors present the relation between Section 5.3 and 5.4 in a clearer way. It's confusing to read how one can perturb a prediction in Section 5.3 before he/she read the algorithm in Section 5.4. I suggest put the algorithm along with the perturbation-driven paradigm at the beginning of Section 5.

A: Thanks for catching the ordering problem and providing a soluation. To address the problem, we add references to the optimization discussed in section 5.4 in the beginning of section 5 as well as in section 5.3. We didn't move all the optimization discussion in the beginning of the section 5, because we believe that presenting details of the all three perturbations (input, attention, prediciton) without any discussion about the interface can also be confusing. And the separation between the operation and UI componenet may also make the description hard to read (requre the reader to refer back to the computation discussion while describing the user interface)

@@

Q8, Except for the presentation issue, the contribution of this paper is somewhat limited to only NLI models with attention mechanisms. It's not sure how the visualization can be easily extended for other NLP tasks. For example, I am not sure how the proposed visualization can easily adapt to machine translation, considering that there are no neutral/contradiction/entailment labels.

A: The composability of the system is the key to extend its capability. The proposed API allow user to include only the panel that is necessary for the given task. For new task (i.e., machine translation), we do not need to include the prediction view for NLI task. We can reuse the attention view to representating the alignment between the original sentence and the translated sentence and introduce new panel for visualizing other information that may not present in the NLI model. However, the main idea is to allow use the combine the panels/functionality to suit their goals.

@@

Q9, In addition, the perturbation-driven scheme is somewhat related to the gradient-based analysis (both are the local analysis of the model on a specific prediction). It is better if the authors discuss the two and differentiate them.

A: Gradient-based sensitivity analysis can often be carried out for domain that is differentiable (i.e., image pixel values), however, for natural language sentences these techniques often can not directly be adopted directly. [?? add the discussion to where?]

@@

Q10, The evaluation is acceptable but a bit weak. There is no description of what questions have the authors asked during the evaluation sessions. It would be better to have formal interviews and ask the participants to perform specific tasks and observe how they performed.

A: We didn't design a specific list tasks and answers due to the exploratory nature of system as well as the open ended analysis goal. If the user didn't follow the particular exploration path the designer of the question goes, it may find some unexpected result that may not directly corresponds to the "correct" answer, but view the problem from a different perspective. However, to properly evaluation the proposed technique is extremely important, so we plan to recurite more domain experts to test and evaluation the system in the new future.

@@

Q11, Please also proofread the work and fix typos. For example, the word 'work' in Section 3 should be uncountable; In Section 5.3, "the three points defines a triangle" -> "the three points defining a triangle"

A: Thanks for pointing out the grammar issues, we have fixed the mentioned problem. And the renew version has been through a thorough copy edit pass by an editor.

### 2018-06-18 paper/evaluation.tex
update response to review 2

\section{Evaluation and Feedback}
The motivation for designing the proposed tool is the error analysis challenges faced by our long-term collaborators working on natural language inference research. During the entire design and development process, we worked closely with two NLP experts, and their constant evaluation and feedback have helped shape the tool we see today. 
%
Since these two NLP experts have been heavily involved in the design process, they may not be the best candidates for identifying potential issues of the tool due to familiarity.  To uncover the limitation and identify areas for improvement, we gathered a wider audience from either visualization (4 researchers, including 3 Ph.D.-level students and one postdoc researcher) or NLP backgrounds (5 Ph.D.-level students who are familiar with the concept of natural language inference and attention) for obtaining feedback for the tool.

>>>>

\section{Evaluation and Feedback}
%The motivation for designing the proposed tool is the error analysis challenges faced by our long-term collaborators working on natural language inference research. During the entire design and development process, we worked closely with two NLP experts, and their constant evaluation and feedback have helped shape the tool we see today. 
%
As discussed previously, we have worked closely with NLP experts during the development of the tool. However, since these two NLP experts have been heavily involved in the design process, they may not be the best candidates for identifying potential issues of the tool due to familiarity.  To uncover the limitation and identify areas for improvement, we gathered a wider audience from either visualization (4 researchers, including 3 Ph.D.-level students and one postdoc researcher) or NLP backgrounds (5 Ph.D.-level students who are familiar with the concept of natural language inference and attention) for obtaining feedback for the tool.

### 2018-06-18 paper/overview.tex
update response to review 2

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze''; see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-T3}).
%
As discussed in the previous section, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affects other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as a guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enable the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences) inside the model, and the \emph{perturbation} of the prediction (i.e., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, the sentence view  (Section~\ref{sec:sentence}), the attention view (Section~\ref{sec:attention}), the prediction view (Section~\ref{sec:prediction}), the pipeline view (Section~\ref{sec:pipeline}), and the pair summary view (Section~\ref{sec:allPairs}).

>>>>

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze''; see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-T3}).
%
As discussed in the previous section, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affects other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as a guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enable the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences) inside the model, and the \emph{perturbation} of the prediction (i.e., adjust the prediction by making updates to the model, the optimization is disucssed in Section~\ref{sec:pipeline}).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, the sentence view  (Section~\ref{sec:sentence}), the attention view (Section~\ref{sec:attention}), the prediction view (Section~\ref{sec:prediction}), the pipeline view (Section~\ref{sec:pipeline}), and the pair summary view (Section~\ref{sec:allPairs}).

### 2018-06-18 paper/prediction.tex
update response to review 2

To perturb the prediction (\textbf{T3}), we need a way to communicate the reassignment of the predicted label. As illustrated in Fig.~\ref{fig:predictionView}(b), we integrate such an operation in the prediction view. When pressing and dragging the prediction (represented as a circle), the user is presented with the three options (E, N, C) corresponding to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label. The reassignment is applied when the user releases the mouse while hovering on the label of choice.

>>>>

To perturb the prediction (\textbf{T3}) (optimization for solving the prediction perturbation is discussed in Section~\ref{sec:pipeline}), we need a way to communicate the reassignment of the predicted label. As illustrated in Fig.~\ref{fig:predictionView}(b), we integrate such an operation in the prediction view. When pressing and dragging the prediction (represented as a circle), the user is presented with the three options (E, N, C) corresponding to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label. The reassignment is applied when the user releases the mouse while hovering on the label of choice.

### 2018-06-18 paper/relatedWorks.tex
update response to review 2

\textbf{Model Interpretability.}
Before one can solve the model interpretability challenge, it worth to understand what constitutes an explanation for a machine learning model.
As discussed in~\cite{Lipton2016, Doshi-Velez2017}, there is little consensus on what interpretability is in the context of machine learning. A taxonomy is suggested~\cite{Doshi-Velez2017} to promote a more rigorous study on such a subject. In the work by Lipton~\cite{Lipton2016}, many philosophical questions related to model interpretability are examined and the study also challenged the commonly held belief that linear models are more interpretable than deep neural networks.
%
In the machine learning community, one of the most studied areas in model interpretability is the interpretation of the convolution neural network (CNN) ~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, OlahSatyanarayanJohnson2018}.
%
These methods either identify the pattern that maximally activates a part of the network (referred to as \emph{feature visualization}, i.e., what feature does the given neuron, channel, or layer captures?), or reveals what areas in the input that contribute most to the predicted label (often referred to as \emph{attribution}, i.e., attribute the prediction label in the input domain).
%
Besides the techniques designed for specific types of models, a few works~\cite{RibeiroSinghGuestrin2016, KrausePererNg2016, LundbergLee2017} approach the interpretability challenge from a model agnostic perspective, which makes them applicable to different applications. In the LIME~\cite{RibeiroSinghGuestrin2016} work, a simpler linear model is fitted near the prediction of interests to provide a simpler but easier to understand snapshot of the classifier near a specific input.
%
Despite being invaluable for generating a human understandable explanation, however, the lack of the ability to access and explore the internal states of the system limits their application. 
%which is vitally important for making sense of the inner works of a model, 
To avoid similar limitations, the proposed system enables direct access and manipulation of the internals of the model.

>>>>

\textbf{Model Interpretability.}
Before one can solve the model interpretability challenge, it worth to understand what constitutes an explanation for a machine learning model.
As discussed in~\cite{Lipton2016, Doshi-Velez2017}, there is little consensus on what interpretability is in the context of machine learning. A taxonomy is suggested~\cite{Doshi-Velez2017} to promote a more rigorous study on such a subject. In the work by Lipton~\cite{Lipton2016}, many philosophical questions related to model interpretability are examined and the study also challenged the commonly held belief that linear models are more interpretable than deep neural networks.
%
In the machine learning community, one of the most studied areas in model interpretability is the interpretation of the convolution neural network (CNN) ~\cite{SimonyanVedaldiZisserman2013, ZeilerFergus2014, YosinskiCluneNguyen2015, OlahMordvintsevSchubert2017, OlahSatyanarayanJohnson2018}.
%
These methods either identify the pattern that maximally activates a part of the network (referred to as \emph{feature visualization}, i.e., what feature does the given neuron, channel, or layer captures?), or reveals what areas in the input that contribute most to the predicted label (often referred to as \emph{attribution}, i.e., attribute the prediction label in the input domain).
%
Besides the techniques designed for specific types of models, a few work~\cite{RibeiroSinghGuestrin2016, KrausePererNg2016, LundbergLee2017} approach the interpretability challenge from a model agnostic perspective, which makes them applicable to different applications. In the LIME~\cite{RibeiroSinghGuestrin2016} work, a simpler linear model is fitted near the prediction of interests to provide a simpler but easier to understand snapshot of the classifier near a specific input.
%
Despite being invaluable for generating a human understandable explanation, however, the lack of the ability to access and explore the internal states of the system limits their application.
%which is vitally important for making sense of the inner works of a model,
To avoid similar limitations, the proposed system enables direct access and manipulation of the internals of the model.

### 2018-06-18 paper/task.tex
update response to review 2

\section{Task Analysis}
\label{sec:task}
Identifying the analysis goals and the corresponding visualization tasks is the first significant step in designing a useful tool. In this section, we examine how domain experts normally approach the task of analyzing a model (i.e., assess its strength and weakness) and identify where and how an interactive visualization system can aid in such a process.
Then, we distill the numerous discussions with domain experts into concrete tasks, which guide the design and implementation of the proposed tool.

During our long-term collaboration with NLP experts, we have conducted extensive discussions to understand the common approach employed by researcher for assessing the behavior of a model.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.

>>>>

\section{Task Analysis}
\label{sec:task}
Identifying the analysis goals and the corresponding visualization tasks is the first significant step in designing a useful tool. In this section, we examine how domain experts normally approach the task of analyzing a model (i.e., assess its strength and weakness) and identify where and how an interactive visualization system can aid in such a process.

\shusen{
The primary motivation for designing the proposed tool is the error analysis challenges faced by our long-term collaborators working on natural language inference research. During the entire design and development process, we worked closely with two NLP experts via weekly meetings in a period of roughly six month (however, our previous collaborations extends well beyond this period), their constant evaluation and feedback have helped shape the tool we see today.
}
%Then, we distill the numerous discussions with domain experts into concrete tasks, which guide the design and implementation of the proposed tool.
During our long-term collaboration with two NLP experts, we have conducted extensive discussions to understand the common approach employed by researchers for assessing the behavior of a model.% as well as provide feedback on the various prototyping stages of the proposed visualization.%
The predictive accuracy has its place as an objective evaluation metric to measure the overall effectiveness of the model; however, the accuracy number alone does not provide the full story.
For example, the model may produce correct predictions for the ``wrong'' reasons (e.g., pick up an unintended pattern in the training data that cannot be generalized in real-word scenarios).
%
As a result, the domain experts rely on an exploration-centric approach to conduct error analysis and obtain intuition.
